---
title: "Tips - DataFrame"
author: "Shunsuke"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_section: true
    theme: spacelab
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, prompt=TRUE)
```



# What is Data Frame in R?

Data frame is a data structure to make it easy to handle two dimensional data. 

Strictly speaking, data frame is a **'list'** with an attribute the name and value of which are **"class"** and **"data.frame"**. However, it can be handled like a 'matrix' by pre-defined generic functions. 

```{r}
df <- data.frame(x=7:10, y=c("a", "b", "c", "d"))
df
typeof(df)
class(df)
attributes(df)
```

> Data frame is a two dimensional data structure in R. It is a special case of a list which has each component of equal length.

https://www.datamentor.io/r-programming/data-frame/

> Data frames are the R structures which most closely mimic the SAS or SPSS data set, i.e. a “cases by variables” matrix of data.
> 
> A data frame is a list of vectors, factors, and/or matrices all having the same length (number of rows in the case of matrices). In addition, a data frame generally has a names attribute labeling the variables and a row.names attribute for labeling the cases. 

https://cran.r-project.org/doc/manuals/r-release/R-lang.html#Data-frame-objects



----

# Functions & syntax sugar

## Create Data Frame

Use data.frame()

```{r}
d0 <- data.frame(x=7:10, y=c("a", "b", "a", "c")); d0

# Row names can be speficied by the argument of "row.names".
d1 <- data.frame(x=7:10, y=c("a", "b", "a", "c"), row.names = c("r1", "r2", "r3", "r4")); d1

# Vectors of 'string' are automatically converted to Factor
d1$y
# This automatic conversion can be stopped by the argument of "stringsAsFactors".
d2 <- data.frame(x=7:10, y=c("a", "b", "a", "c"), stringsAsFactors = FALSE)
d2$y
```

## Get elements

> We can use either **[**, **[[** or **$** operator to access columns of data frame.
> 
> Accessing with **[[** or **$** is similar. However, it differs for **[** in that, indexing with **[** will return us a data frame but the other two will reduce it into a vector.

https://www.datamentor.io/r-programming/data-frame/


```{r}
d3 <- data.frame(x=7:10, y=c("a", "b", "a", "c"), z=c(0,1.1,2,3.5)); d3
d3[1]; class(d3[1])
d3[[1]]; typeof(d3[[1]])
d3$z; typeof(d3$z)
```

> Data frames can be accessed like a matrix by providing index for row and column.

https://www.datamentor.io/r-programming/data-frame/

```{r}
d3[1:2]     # get the first and second columns
d3[1:2,]    # get the first and second rows
d3[1:3, 1:2]
d3[1:2, c('x','z')]
```


## Access to properties {.tabset}

* nrow()
* ncol()
* dim()
* typeof()
* rownames()
* colmatnames()



### Demension or type

```{r}
nrow(d3)
ncol(d3)
dim(d3)
typeof(d3)
```

### Label

```{r}
rownames(d3) <- c("r1", "r2", "r3", "r4")
rownames(d3)
colnames(d3) <- c("c1", "c2", "c3")
colnames(d3)

# Specify rows and columns by using labels
d3["r2","c3"]
d3[c("r1", "r3"),]
```




## Add columns or rows

You can use cbind() or rbind().

```{r}
d <- data.frame(x=7:10, y=c("a", "b", "c", "d"))
cbind(d, newcol=c("Z", "Y", "X", "V"))
rbind(d, list(100, "X"))
```

## Add columns (destructive)

There are two methods: $ and brackets.

```{r}
d$nc1 = 1:4
d['nc2'] = c(9,7,5,3)
d
```








