---
title: "Tips - Vector"
author: "Shunsuke"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_section: true
    theme: spacelab
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, prompt=TRUE)
```

# What is vector in R?

> Vector is a basic data structure in R. It contains element of the same type. The data types can be **logical, integer, double, character, complex or raw**.
>
> A vector’s type can be checked with the typeof() function.
>
> Another important property of a vector is its length. This is the number of elements in the vector and can be checked with the function length().

https://www.datamentor.io/r-programming/vector/

Other important points are listed below. 

* Vector containing only one element can be defined. 
* Vector can include not only numbers but also booleans and strings.
* However, if vector include strings, arithmetic operations cannot be applied to it. 
* If arithmetic operations are applied to vector, TRUE and FALSE are automatically converted to 1 and 0 respectively. 

```{r, collapse=TRUE}
is.vector(1:3)
is.vector( c(TRUE, FALSE) )
is.vector("str")
```

```{r, collapse=TRUE}
c(TRUE, FALSE) * 10
```


----

# Functions & syntax sugar

## Create a vector {.tabset}

**List of functions**

* c()
* 1:10
* seq()
* rep()

**Examples**

### Simplest ways

```{r, collapse=TRUE}
c(-4 , TRUE, "a", FALSE, 5, "bbbb")
1:10*3 + 1
```

### Useful functions

```{r}
seq(1, 20, by=5)         # 等差数列   1 6 11 16
seq(1,5, length.out=9)   # devide the range from 1 to 5 by N-1 (N=9 in this case). 
rep(1:3, times=3)        # repeat c(1,2,3) three times
rep(1:3, each=3)         # repeat each element of c(1,2,3)
```


## Access to properties {.tabset}

* length()
* typeof()
* names()

**Examples**

### Get length or type

```{r, collapse=TRUE}
# Get the length of a vector
length(1:4)

# Get the type of elements
typeof(3:5)
typeof(c(TRUE, FALSE))
typeof(c("a", "bb", 3, TRUE))
```


### Access to labels

```{r, collapse=TRUE}
v <- c(3, 5, 8)
names(v) <- c("x", "y", "z")  # set labels
names(v)                      # get labels
v[ c("x","z") ]
```

## Get elements {.tabset}

* x[]
* head()
* tail()

**Examples**

```{r}
# Prepare a vector.
x <- c(10, 20, 30, 40, 50, 60)
```

### Included elements

```{r}
# Specify included elements. 
x[3:5]                                   # use a logical vector as index
x[3:length(x)]
x[c(1,3,4)]
x[c(TRUE,FALSE,TRUE,FALSE,FALSE,FALSE)]  # use a logical vector as index
x[x %% 3 == 0]                           # specify conditions
```

### Excluded elements

```{r}
# Specify excluded elements
x[-1]
x[-2:-4]
x[c(-2,-5)]
```

### head(), tail()

```{r}
head(x, 3)
tail(x, 2)            # 50 60   (指定した数の要素を末尾から)
```


## Get indexes

Function to get indexes satisfing given conditions:

* which()

**Examples**

```{r}
which(x > 20)         # indexを取得したい時
which(c(TRUE, FALSE, TRUE))
```

## Edit vectors {.tabset}

* c()
* append()
* replace()
* rev()
* sort()
* unique()

**Examples**

### Concatenate

```{r}
x1 <- c(1, 2)
x2 <- c(30, 40, 50)
n <- 600
c(x1, x2, n, 7000)
append(x1, x2)            # (単純に連結)
append(x1, x2, after=1)   # (ベクトルの途中に値を挿入)
```

### Change values

```{r}
z <- c(10, 20, 30, 40, 50, 60)
replace( z, 2:4, 0 )
replace( z, 2:4, c(2,3,4) )
replace( z, 1:4, c(11,22) )
replace( z, z<50, c(11, 12, 13, 14) )
```

### Change order

```{r}
# reverse
rev( c(10, 20, 30) )
# sort
sort( c(3,5,2,1,4) )
sort( c(3,5,2,1,4), decreasing=TRUE )
```

### Remove duplicate elements

```{r}
unique( c(10:5, 6:17))
```


## Edit vectors (destructive) {.tabset}

```{r}
z <- c(10, 20, 30, 40, 50, 60)
z[2:4] <- 0 ; z
z[2:4] <- c(2,3,4) ; z
z[1:4] <- c(1,2); z
z[z<50] <- c(11, 12, 13, 14) ; z
```


----

# Manipulate types of elements

**Basic types**

* is.logical(), as.logical()
* is.numeric(), as.numeric()
* is.double(), as.double()
* is.character(), as.character()

**null-able**

* is.null()
* is.na()
* is.nan()
* is.finite()
* is.infinite()

----

# Operators

## Algebraic calculation {.tabset}

* +, -, *, /
* %*%
* %o%
* expand.grid()

### Addition, Multiplication

```{r}
# Addition
1:3 + c(10, 20, 30)
# Scalar multiplication 
100 * 4:6
```

### Inner product

$a \cdot b = \sum a_i b_i$

```{r}
# Inner product
mat <- 1:3 %*% c(1, 10, 100); mat    # 1x1 matrix is returned.
drop(mat)                            # conversion from matrix to vector. 
```

### Outer (Tensor) product

$a \otimes b = \sum \sum a_i b_j$

```{r}
# Outer (Tensor) product (テンソル積)
1:3 %o% c(1, 10, 100)

#matrix(1:9, 3) %o% c(1, 10, 100)
```

### 直積

```{r}
expand.grid(1:3, 8:9)
```



## Conditional Binary {.tabset}

* ==
* <, >, >=, <=
* %in%
* !
* &, |
* &&, ||

Operators for conditional returns a logical vector except for '&&' and '||'.

### Returning a vector

```{r}
x <- c(10, 20, 30, 40, 50, 60)

x == 30
x >= 30
x %in% 22:45     # Does each element of x matches values included in 22:45?

!c(TRUE, FALSE, FALSE)
(x > 11) & (x < 43)
```

### && and ||

'&&', and '||' only check the first element of vectors and return a scalar.

```{r}
c(TRUE, TRUE, FALSE) && c(TRUE, FALSE)
c(FALSE, TRUE, TRUE) && c(TRUE, FALSE)
c(FALSE, TRUE) || c(FALSE, TRUE)
```



----

# Utilities

## Statistical Properties {.tabset}

* mean()
* median()
* var()
* sd()
* max(), cummax()
* min(), cummin()
* sum(), cumsum()
* prod(), cumprod()

A function for 'mode' is not prepared.

**Examples**

```{r}
X <- c(1,1,2,3,3,3,7,4,5,0)
```

### Return scalar

```{r}
mean(X)
median(X)
var(X)    # Variance (the denominator 'n-1' is used)
sd(X)     # Standard deviation (the denominator 'n-1' is used)
max(X)
min(X)
sum(X)
prod(X)
```

### Return vector

```{r}
# Cumulative properties
cummax(X)
cummin(X)
cumsum(X)
cumprod(X)
```


## High-order functions

* lapply(X, FUN, )
* sapply(X, FUN, simplify = TRUE, USE.NAMES = TRUE)

```{r}
sapply(1:10, function(x) {x * 2})
sapply(1:10, function(x) {1:x})
lapply(1:10, function(x) {x * 2})  # return vector (or matrix/array) if possible.  
lapply(1:10, function(x) {1:x})    # return list if impossible.
```


## For logical vectors

* all()
* any()

```{r}
all(c(TRUE, TRUE, FALSE))
all(c(TRUE, TRUE, TRUE))
any(c(TRUE, TRUE, FALSE))
any(c(FALSE, FALSE, FALSE))
```
