---
title: "Tips_Matrix"
author: "Shunsuke"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_section: true
    theme: spacelab
    highlight: zenburn
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, prompt=TRUE)
```

# What is matrix in R?

> Matrix is a two dimensional data structure in R programming.
> Matrix is similar to vector but additionally contains the dimension attribute. All attributes of an object can be checked with the **attributes()** function (dimension can be checked directly with the dim() function).
> We can check if a variable is a matrix or not with the **class()** function.

https://www.datamentor.io/r-programming/matrix/

Other important characters are listed below.

* As with vector, all elements have a same data type
* The data type can be checked by **typeof()**

```{r, collapse=TRUE}
m <- matrix(1:12, 3)
attributes(m)
class(m)
typeof(m)
c(is.matrix(m), is.vector(m))
c(is.integer(m), is.logical(m), is.double(m))
```


----

# Functions & syntax sugar

## Create a matrix

* matrix()
* cbind()
* rbind()

```{r}
matrix(1:12, 3)
matrix(1:12, ncol=4)
cbind(1:3, 4:6, 7:9, 10:12)
rbind(c(1,4,7,10), c(2,5,8,11), c(3,6,9,12))
```

## Access to properties

* ncol()
* nrow()
* dim()
* typeof()
* rownames()
* colmatnames()


## Get elements

* mat[row,col]
* mat[row,]
* mat[,col]

```{r}
mat <- matrix(1:12, 3); mat
mat[2,3]
mat[1,]
mat[c(1,3),]
mat[2:nrow(mat),]
mat[, 1:3]          # (1列目 ～ 3列目)
mat[,2:ncol(mat)]   # (2列目から最後の列まで)
mat[2:3, c(2,4)]    # (2行目・3行目 の 2列目と4列目の要素)
```



