---
title: "Tips_Matrix"
author: "Shunsuke"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_section: true
    theme: spacelab
    highlight: zenburn
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, prompt=TRUE)
```

# What is matrix in R?

> Matrix is a two dimensional data structure in R programming.
> Matrix is similar to vector but additionally contains the dimension attribute. All attributes of an object can be checked with the **attributes()** function (dimension can be checked directly with the dim() function).
> We can check if a variable is a matrix or not with the **class()** function.

https://www.datamentor.io/r-programming/matrix/

Other important characters are listed below.

* As with vector, all elements have a same data type
* The data type can be checked by **typeof()**

```{r, collapse=TRUE}
m <- matrix(1:12, 3)
attributes(m)
class(m)
typeof(m)
c(is.matrix(m), is.vector(m))
c(is.integer(m), is.logical(m), is.double(m))
```


----

# Functions & syntax sugar

## Create a matrix {.tabset}

* matrix()
* cbind()
* rbind()
* diag()

### Simplest ways

```{r}
matrix(1:12, 3)
matrix(1:12, ncol=4)
matrix(1:12, 3, byrow=TRUE)
cbind(1:3, 4:6, 7:9, 10:12)
rbind(c(1,4,7,10), c(2,5,8,11), c(3,6,9,12))
```

### Useful functions

```{r}
diag(3)    # 3 x 3 Identity matrix
diag(1:3)  # 3 x 3 diagonal matrix
```


## Get elements or properties {.tabset}

* m[row,col]
* nrow()
* ncol()
* dim()
* typeof()
* rownames()
* colmatnames()


```{r}
m <- matrix(1:12, 3, byrow=TRUE)
```

$m = \begin{pmatrix} 1 & 2 & 3 & 4 \\ 5 & 6 & 7 & 8 \\ 9 & 10 & 11 & 12 \end{pmatrix}$

### Elements

```{r}
# Specify single element
m[2,3]

# Specify row(s)
m[1,]                   # single row (a vector is returned)
m[c(1,3),]
m[2:nrow(m),]
m[c(TRUE,TRUE,FALSE),]

# Specify columns
m[, 1:3]
m[,2:ncol(m)]

# Specify rows and columns
m[2:3, c(2,4)]
```


### Demension or type

```{r}
nrow(m)
ncol(m)
dim(m)
typeof(m)
```

### Label

```{r}
rownames(m) <- c("r1", "r2", "r3")
rownames(m)
colnames(m) <- c("c1", "c2", "c3", "c4")
colnames(m)

# Specify rows and columns by using labels
m["r2","c3"]
m[c("r1", "r3"),]
```


## Edit Matrix {.tabset}

* cbind()
* rbind()
* t()
* nrow():1, ncol():1

```{r}
m1 <- matrix(1:4, 2, byrow=TRUE)
m2 <- matrix(1:4*10, 2, byrow=TRUE)
```

$m1 = \begin{pmatrix} 1 & 2 \\ 3 & 4 \end{pmatrix}, \ \ m2 = \begin{pmatrix} 10 & 20 \\ 30 & 40 \end{pmatrix}$

### Concatenate

```{r}
cbind(m1, m2)
cbind(m1, c(100,200))
rbind(m1, m2)
rbind(m1, c(100,200))
```

### Change order

```{r}
t(m1)
m1[nrow(m1):1,]
m1[,ncol(m1):1]
```

----

# Algebraic calculation

```{r}
A <- matrix(1:4, 2, byrow=TRUE)
B <- matrix(1:4*10, 2, byrow=TRUE)
```

$A = \begin{pmatrix} 1 & 2 \\ 3 & 4 \end{pmatrix}, \ \ B = \begin{pmatrix} 10 & 20 \\ 30 & 40 \end{pmatrix}$

## Basic

```{r}
# Addition
A + B
# Scalar multiplication 
100 * A
# Matrix multiplication
A %*% B
```

$A B = \begin{pmatrix} 1 & 2 \\ 3 & 4 \end{pmatrix} \begin{pmatrix} 10 & 20 \\ 30 & 40 \end{pmatrix} = \begin{pmatrix} 10 + 60 & 20 + 80 \\ 30 + 120 & 60 + 160 \end{pmatrix}$

----


## Determinant

* det()
* solve()

```{r}
det(A)           # Determinant
solve(A)         # Inverse of A
A %*% solve(A)   # (small error may be included)
```


$$ \det{(A)} = (1 \cdot 4) - (2 \cdot 3) = -2 $$
