---
title: "Tips - Vector"
author: "Shunsuke"
date: "2019年6月9日"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    number_section: true
    theme: spacelab
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, prompt=TRUE)
```

# What is vector in R?

> Vector is a basic data structure in R. It contains element of the same type. The data types can be logical, integer, double, character, complex or raw.
>
> A vector’s type can be checked with the typeof() function.
>
> Another important property of a vector is its length. This is the number of elements in the vector and can be checked with the function length().

https://www.datamentor.io/r-programming/vector/

Other important points are listed below. 

* Vector containing only one element can be defined. 
* Vector can include not only numbers but also booleans and strings.
* However, if vector include strings, arithmetic operations cannot be applied to it. 
* If arithmetic operations are applied to vector, TRUE and FALSE are automatically converted to 1 and 0 respectively. 

```{r, collapse=TRUE}
is.vector(1:3)
is.vector( c(TRUE, FALSE) )
is.vector("str")
```

```{r, collapse=TRUE}
c(TRUE, FALSE) * 10
```


----

# Functions & syntax sugar

## Create a vector

* c()
* seq()
* rep()
* 1:10

**The simplest way to a create vector**

```{r, collapse=TRUE}
c(-4 , TRUE, "a", FALSE, 5, "bbbb")
```

**Create vectors consisting of numbers. **

```{r}
1:10*3 + 1
seq(1, 20, by=5)         # 等差数列   1 6 11 16
seq(1,5, length.out=9)   # devide the range from 1 to 5 by N-1 (N=9 in this case). 
rep(1:3, times=3)        # repeat c(1,2,3) three times
rep(1:3, each=3)         # repeat each element of c(1,2,3)
```


## Access to properties 

* length()
* typeof()
* names()

**Get type or length**

```{r, collapse=TRUE}
# Get the length of a vector
length(1:4)

# Get the type of elements
typeof(3:5)
typeof(c(TRUE, FALSE))
typeof(c("a", "bb", 3, TRUE))
```


**Access to labels**

```{r, collapse=TRUE}
v <- c(3, 5, 8)
names(v) <- c("x", "y", "z")  # set labels
names(v)                      # get labels
v[ c("x","z") ]
```

## Get elements

* x[]
* tail()

**specify indexes**

```{r}
x <- c(10, 20, 30, 40, 50, 60)

# Specify required elements
x[3:5]                                   # use a logical vector as index
x[3:length(x)]
x[c(1,3,4)]
x[c(TRUE,FALSE,TRUE,FALSE,FALSE,FALSE)]  # use a logical vector as index
x[x %% 3 == 0]                           # specify conditions

# Use tail() function
tail(x, 2)            # 50 60   (指定した数の要素を末尾から)

# Specify excluded elements
x[-1]
x[-2:-4]
x[c(-2,-5)]
```


## Get indexes satisfing given conditions

* which()

```{r}
which(x > 20)         # indexを取得したい時
```


## Concatenate





# Utilities

* cumsum()

