---
title: "Usage of defaultSetting.R"
author: "Shunsuke"
date: '`r format(Sys.time(), "%y/%m/%d %H:%M")`'
output:
  pdf_document:
    highlight: tango
    latex_engine: lualatex
    number_section: yes
  html_document:
    code_folding: hide
    highlight: zenburn
    number_section: yes
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
header-includes: \setmainfont{Arial}
---


<style type="text/css">

body{ /* Normal  */
  font-size: 16px;
  #font-family: Verdana, Helvetica, Arial, sans-serif;
  #color: #456788;
  color: "#090909";
}

h1 {
  font-size: 28px;
  font-weight: bold;
  color: "#030303";
}
h2 {
  font-size: 24px;
  font-weight: bold;
  font-color: "#030303";
}

code.r {
  font-size: 16px;
  #background-color: #eeeeee;
} 

</style>


\fontsize{12}{16} 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, prompt=TRUE, comment="#")
source("../defaultSetting.R")
vec <- 1:10 * 10
mat <- rbind( 11:13, 21:23, 31:33 )
df <- data.frame(A = 1:5, B=c(3, -3, 0, 3, 1), C=c(99,88,77,66,55))
```

---

# Data 

## Vector


* v.indexClosestValue(vec, x)
* v.closestValue(vec, x)

```{r collapse=TRUE}
vec <- 1:10 * 10; vec
v.indexClosestValue(vec, 31)
v.closestValue(vec, 75)
```


## Matrix

* m.rrev(mat)
* m.crev(mat)

```{r collapse=TRUE, prompt=TRUE}
mat <- rbind( 11:13, 21:23, 31:33 ); mat
m.rrev( mat )
m.crev( mat )
```

## Data Frame

* df.orderBy(df, colname, decreasing=FALSE)

```{r df, collapse=TRUE, prompt=TRUE}
df_ <- data.frame(A = 1:5, B=c(3, -3, 0, 3, 1), C=c(99,88,77,66,55)); df_
df.orderBy(df_, "B")
df.orderBy(df_, 3)
df.orderBy(df, "B", decreasing=TRUE)
```

## Probability

* cum.probability(values, decreasing=FALSE)

```{r, collapse=TRUE}
# If you want to create a graph of cumulative distribution, you should use ecdf() function.
#   fCP <- ecdf( c(2, 1:5) ); plot(fCP)
cum.probability(c(3,76,58,24,100,1,1))
```

---

# IO clipboard

* rcbmat(header, ...)
* wcbmat(data, header, sep="\t", ...)

```{r, eval=FALSE, prompt=FALSE}
# This function is used to import data from a Excel sheet via the clipboard. 
# You need to copy cells on Excel before calling this function. 
> data <- rcbmat(TRUE)    # If the data includes a header, set "TRUE". 

# This function is used to export matrix or data.frame to the clipboard.
# After calling this function, you can paste the data to a Excel sheet.
> wcbmat(data, TRUE)      # If the data includes a header, set "TRUE". 
```


# Log

* FATAL()
* ERROR()
* WARN()
* INFO()
* DEBUG()
* TRACE()
* DEBUG_LOG$FILE
* DEBUG_LOG$LEVEL

    * DEBUG_LOG$LEVEL_WARN
    * DEBUG_LOG$LEVEL_INFO
    * DEBUG_LOG$LEVEL_DEBUG
    * DEBUG_LOG$LEVEL_TRACE

```{r eval=FALSE}
DEBUG_LOG$FILE <- "log.txt"              # speficy name of the log file.
DEBUG_LOG$LEVEL <- DEBUG_LOG$LEVEL_INFO  # speficy the log level.

f <- 3.45; s <- "abc"
INFO("a = %f, s = %s", a, s)             # output a log message.
DEBUG("log message")                     # this message is not outputed.

DEBUG_LOG$LEVEL <- DEBUG_LOG$LEVEL_DEBUG
DEBUG("log message")                     # this message is outputed.
```




---


$$u = \begin{pmatrix} 10 & 20 & 30 & 40 & 50 & 60 & 70 & 80 & 90 & 100 \end{pmatrix}$$

$$\begin{array}{ccc}
x_{11} & x_{12} & x_{13}\\
x_{21} & x_{22} & x_{23}
\end{array}$$


# Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
